A quantitative framework for investigating the reliability of network construction

### 2018-03-20 manuscript/coverletter.tex
coverletter drafted

% The instructions are for a single, 300-word paragraph proposal. The large paragraph is now at 299 words, and we can maybe hope that they won't count the preamble and nice closing.

@@

    My co-authors and I cordially invite you to consider the 
    following proposal for an Ideas \& Perspectives article in 
    \myjournal:

    Representing communities as networks of interactions (links) between taxa (nodes) is a central approach in modern community ecology. These networks are usually treated as fixed, deterministic entities. Yet, researchers working with ecological networks widely acknowledge that empirical data are in fact samples of a community and come with major caveats. The probability of observing different interactions depends on sampling effort/methodology and local conditions as well as the intrinsic probability of any interaction. These contingencies mean that there are many reasons one might fail to observe an interaction (e.g., it might occur sometimes but not during sampling due to weather). Problematically, these sources of non-interaction are often indistinguishable within an interaction matrix. As a potential solution, we propose a Bayesian framework which partitions the uncertainty associated with missing interactions into explicit components and allows us to place confidence intervals around each estimated interaction probability. The Bayesian framework also facilitates the use of prior information (e.g., species traits) to reduce uncertainty around some interactions. We will illustrate this framework using an extraordinarily  well-replicated set of plant-galler-parasitoid networks recently published as a data paper (Kopelke \emph{et al.}, 2017. \textbf{Ecology} 98: 1730). Using this dataset, we will demonstrate how uncertainty surrounding each interaction probability affects network-level metrics. Acknowledging this uncertainty is crucial for valid comparison of networks along gradients. The authors involved in this study contribute a range of valuable perspectives. Dominique Gravel has developed techniques for novel types of interaction distribution modelling. Anna Ekl\"{o}f contributes extensive experience with probabilistic network analyses. Tomas Roslin has vast experience with gathering and using network data and has first-hand experience with each source of variation we describe. The two junior authors, Kate Wootton and Alyssa Cirtwill, work with both empirical and simulated networks and have a strong interest in developing more robust methods for network analyses.

    We are confident that the applicability of this method to all types of ecological networks and the importance of addressing sources of uncertainty in the data will mean that this article is of interest to a broad readership. We hope that you agree and eagerly await your reply.

@@

>>>>

% Borrowing heavily from the proposal

@@

    My co-authors and I are pleased to submit the attached manuscript \emph{A quantitative framework for investigating the reliability of network construction} for consideration as a Novel Methods paper in \emph{Ecology Letters} as per our previous correspondence with Senior Editor John Drake. As instructed, we have submitted the manuscript under the Ideas \& Perspectives mechanism (and following the associated formatting guidelines).

    We begin the manuscript by reviewing the many sources of variation inherent in ecological networks. Some sources of variation can be reduced through improved sampling, but others are intrinsic properties of ecological communities. This uncertainty limits the accuracy of network studies which consider communities as static, invariant entities and demonstrates the need for descriptions of network structure that acknowledge and address the uncertainty surrounding estimates of interspecific interactions.

    As a potential solution, we propose a Bayesian framework which partitions the uncertainty associated with missing interactions into explicit components and allows us to place confidence intervals around each estimated interaction probability. The Bayesian framework also facilitates the use of prior information (e.g., species traits) to reduce uncertainty around some interactions. We illustrate this framework using an extraordinarily  well-replicated set of plant-galler-parasitoid networks recently published as a data paper (Kopelke \emph{et al.}, 2017. \textbf{Ecology} 98: 1730). Using this dataset, we demonstrate how uncertainty surrounding each interaction probability affects network-level metrics. Our derivation of the Bayesian framework and quantitative examples are constructed so as to be accessible to readers with little or no prior experience with Bayesian statistics. We hope that this gentle introduction will help the broadest possible set of readers apply the framework we propose.

    The present work is outstandingly novel relative to the recent work by one of the authors (Tomas Roslin) cited in this manuscript. Kopelke \emph{et al.} (2017) is the data paper describing the \emph{Salix}-galler-parasitoid system used for the quantitative analyses in the present work. Kopelke \emph{et al.} (2017) does not contain any analyses and so does not affect the novelty of the present work in any way.

    We appreciate the chance to submit one of the first Novel Methods papers at \emph{Ecology Letters}. We are confident that the fundamental conceptual shift implied by our approach, accompanied by clear examples and easy-to-use scripts, can achieve the high standard you are looking for. We hope that you agree and eagerly await your reply.

@@

% Suggested reviewers:
